### 2023-11-09 submission/manuscript.Rmd
Fix some typos that appeared in proofs

Confusion over the terms is demonstrated in the choice of citations that McMurdie and Holmes used to define rarefying (i.e., [@NavasMolina2013; @Hughes2005; @Koren2013]). The cited studies either did not use the words "rarefy" or "rarefying" or used them interchangeably with rarefy as a verb form of rarefaction. For example, Hughes and Hellman did not use "rarefy" [@Hughes2005]. Rather they used "rarefaction" in the traditional sense with multiple subsamplings. Meanwhile, the QIIME-based literature appears to use "rarefy" and "rarefaction" interchangeably to mean only a single subsampling [@NavasMolina2013; @Koren2013]. Confusion comes from the WNWN authors' admonition that "[i]n many cases researchers have also failed to repeat the random subsampling step". This seems to call on researchers to use rarefaction rather than the single subsampling step. Subsequent researchers have continued to conflate the terms when citing WNWN (**Supplemental Text**). An exemplar of the confusion is the creation of a technique that uses "repeatedly rarefying" as an approach distinct from rarefaction when they were in fact re-proposing traditional rarefaction [@Cameron2021].

@@

The code contained within `simulation-cluster-accuracy-server.Rmd` was spilt into individual R scripts and modified to make the generation of the simulated communities more modular and scalable. The code for generating the simulated communities was run using the `nr-s1` Conda environment. The ocean and feces parent distributions were generated as described above. Five variables were altered to simulate each set of communities. First, eight mixing fractions were used to manipulate the effect size between the two simulated treatment groups with 40 samples per group. These were the same effect sizes used in the WNWN study (1.00 [i.e., no difference], 1.15, 1.25, 1.50, 1.75, 2.00, 2.50, and 3.50). Second, the number of sequences in each of the 80 samples was either randomly selected from the `r gp_n_samples` library sizes contained within the GlobalPatterns dataset or from a log distribution of 80 sequencing depths evenly distributed between the most shallow (`r gp_min` sequences) and deeply (`r gp_max` sequences) sequenced samples in the GlobalPatterns dataset (Figure 10). Third, the resulting 80 sequencing depths were scaled so that the median sequencing depth across the 80 samples was either 1,000, 2,000, 5,000, 10,000, or 50,000 sequences. Fourth, the sequencing depth of each sample was either randomly assigned to each treatment group or assigned so that those depths less than or greater than the median were assigned to separate treatment groups. Finally, each set of conditions was replicated 100 times. These five parameters resulted in 1,600 simulated datasets (8 effect sizes x 2 sequencing depth models x 5 sequencing depths x 2 sequencing depth assignment models x 100 replicates).

@@

**Figure 3. When the median sequencing depth was 2,000 sequences or more, rarefaction of the entire dataset performed better than removing the smallest 15% of samples when using K-means clustering.** This figure is analogous to Figure 4 except that K-means clustering was used instead of PAM.

**Figure 4. K-means clustering of distances calculated with rarefaction were as good or better than any other normalization method.** This figure is analogous to Figure 3 except that K-means clustering was used instead of PAM, rarefaction on the full dataset was used instead of subsampling to the size of the sample at the 15th percentile, and DESeq Variance Stabilization normalized OTU counts were only used with Euclidean distances.

**Figure 5. Clustering accuracies that used rarefaction were as good or better than the other normalization procedures when a log-scaled distribution of sequencing depths.** This figure is analogous to Figure 9 except that the sequencing depths for each of the 80 samples in each simulation were drawn without replacement from a log-scaled distribution rather than from the GlobalPatterns sequencing depths.

>>>>

Confusion over the terms is demonstrated in the choice of citations that McMurdie and Holmes used to define rarefying (i.e., [@NavasMolina2013; @Hughes2005; @Koren2013]). The cited studies either did not use the words "rarefy" or "rarefying" or used them interchangeably with rarefy as a verb form of rarefaction. For example, Hughes and Hellmann did not use "rarefy" [@Hughes2005]. Rather they used "rarefaction" in the traditional sense with multiple subsamplings. Meanwhile, the QIIME-based literature appears to use "rarefy" and "rarefaction" interchangeably to mean only a single subsampling [@NavasMolina2013; @Koren2013]. Confusion comes from the WNWN authors' admonition that "[i]n many cases researchers have also failed to repeat the random subsampling step". This seems to call on researchers to use rarefaction rather than the single subsampling step. Subsequent researchers have continued to conflate the terms when citing WNWN (**Supplemental Text**). An exemplar of the confusion is the creation of a technique that uses "repeatedly rarefying" as an approach distinct from rarefaction when they were in fact re-proposing traditional rarefaction [@Cameron2021].

@@

The code contained within `simulation-cluster-accuracy-server.Rmd` was spilt into individual R scripts and modified to make the generation of the simulated communities more modular and scalable. The code for generating the simulated communities was run using the `nr-s1` Conda environment. The ocean and feces parent distributions were generated as described above. Five variables were altered to simulate each set of communities. First, eight mixing fractions were used to manipulate the effect size between the two simulated treatment groups with 40 samples per group. These were the same effect sizes used in the WNWN study (1.00 [i.e., no difference], 1.15, 1.25, 1.50, 1.75, 2.00, 2.50, and 3.50). Second, the number of sequences in each of the 80 samples was either randomly selected from the `r gp_n_samples` library sizes contained within the GlobalPatterns dataset or from a log distribution of 80 sequencing depths evenly distributed between the most shallow (`r gp_min` sequences) and deeply (`r gp_max` sequences) sequenced samples in the GlobalPatterns dataset (Figure S6). Third, the resulting 80 sequencing depths were scaled so that the median sequencing depth across the 80 samples was either 1,000, 2,000, 5,000, 10,000, or 50,000 sequences. Fourth, the sequencing depth of each sample was either randomly assigned to each treatment group or assigned so that those depths less than or greater than the median were assigned to separate treatment groups. Finally, each set of conditions was replicated 100 times. These five parameters resulted in 1,600 simulated datasets (8 effect sizes x 2 sequencing depth models x 5 sequencing depths x 2 sequencing depth assignment models x 100 replicates).

@@

**Figure 3. When the median sequencing depth was 2,000 sequences or more, rarefaction of the entire dataset performed better than removing the smallest 15% of samples when using K-means clustering.** This figure is analogous to Figure S4 except that K-means clustering was used instead of PAM.

**Figure 4. K-means clustering of distances calculated with rarefaction were as good or better than any other normalization method.** This figure is analogous to Figure S3 except that K-means clustering was used instead of PAM, rarefaction on the full dataset was used instead of subsampling to the size of the sample at the 15th percentile, and DESeq Variance Stabilization normalized OTU counts were only used with Euclidean distances.

**Figure 5. Clustering accuracies that used rarefaction were as good or better than the other normalization procedures when a log-scaled distribution of sequencing depths.** This figure is analogous to Figure 4 except that the sequencing depths for each of the 80 samples in each simulation were drawn without replacement from a log-scaled distribution rather than from the GlobalPatterns sequencing depths.

### 2023-11-09 submission/manuscript.tex
Fix some typos that appeared in proofs

Confusion over the terms is demonstrated in the choice of citations that
McMurdie and Holmes used to define rarefying (i.e., (13, 27, 28)). The
cited studies either did not use the words ``rarefy'' or ``rarefying''
or used them interchangeably with rarefy as a verb form of rarefaction.
For example, Hughes and Hellman did not use ``rarefy'' (13). Rather they
used ``rarefaction'' in the traditional sense with multiple
subsamplings. Meanwhile, the QIIME-based literature appears to use
``rarefy'' and ``rarefaction'' interchangeably to mean only a single
subsampling (27, 28). Confusion comes from the WNWN authors' admonition
that ``{[}i{]}n many cases researchers have also failed to repeat the
random subsampling step''. This seems to call on researchers to use
rarefaction rather than the single subsampling step. Subsequent
researchers have continued to conflate the terms when citing WNWN
(\textbf{Supplemental Text}). An exemplar of the confusion is the
creation of a technique that uses ``repeatedly rarefying'' as an
approach distinct from rarefaction when they were in fact re-proposing
traditional rarefaction (29).

@@

The code contained within
\texttt{simulation-cluster-accuracy-server.Rmd} was spilt into
individual R scripts and modified to make the generation of the
simulated communities more modular and scalable. The code for generating
the simulated communities was run using the \texttt{nr-s1} Conda
environment. The ocean and feces parent distributions were generated as
described above. Five variables were altered to simulate each set of
communities. First, eight mixing fractions were used to manipulate the
effect size between the two simulated treatment groups with 40 samples
per group. These were the same effect sizes used in the WNWN study (1.00
{[}i.e., no difference{]}, 1.15, 1.25, 1.50, 1.75, 2.00, 2.50, and
3.50). Second, the number of sequences in each of the 80 samples was
either randomly selected from the 26 library sizes contained within the
GlobalPatterns dataset or from a log distribution of 80 sequencing
depths evenly distributed between the most shallow (58,688 sequences)
and deeply (2,357,181 sequences) sequenced samples in the GlobalPatterns
dataset (Figure 10). Third, the resulting 80 sequencing depths were
scaled so that the median sequencing depth across the 80 samples was
either 1,000, 2,000, 5,000, 10,000, or 50,000 sequences. Fourth, the
sequencing depth of each sample was either randomly assigned to each
treatment group or assigned so that those depths less than or greater
than the median were assigned to separate treatment groups. Finally,
each set of conditions was replicated 100 times. These five parameters
resulted in 1,600 simulated datasets (8 effect sizes x 2 sequencing
depth models x 5 sequencing depths x 2 sequencing depth assignment
models x 100 replicates).

@@

\textbf{Figure 3. When the median sequencing depth was 2,000 sequences
or more, rarefaction of the entire dataset performed better than
removing the smallest 15\% of samples when using K-means clustering.}
This figure is analogous to Figure 4 except that K-means clustering was
used instead of PAM.

\textbf{Figure 4. K-means clustering of distances calculated with
rarefaction were as good or better than any other normalization method.}
This figure is analogous to Figure 3 except that K-means clustering was
used instead of PAM, rarefaction on the full dataset was used instead of
subsampling to the size of the sample at the 15th percentile, and DESeq
Variance Stabilization normalized OTU counts were only used with
Euclidean distances.

\textbf{Figure 5. Clustering accuracies that used rarefaction were as
good or better than the other normalization procedures when a log-scaled
distribution of sequencing depths.} This figure is analogous to Figure 9
except that the sequencing depths for each of the 80 samples in each
simulation were drawn without replacement from a log-scaled distribution
rather than from the GlobalPatterns sequencing depths.

>>>>

Confusion over the terms is demonstrated in the choice of citations that
McMurdie and Holmes used to define rarefying (i.e., (13, 27, 28)). The
cited studies either did not use the words ``rarefy'' or ``rarefying''
or used them interchangeably with rarefy as a verb form of rarefaction.
For example, Hughes and Hellmann did not use ``rarefy'' (13). Rather
they used ``rarefaction'' in the traditional sense with multiple
subsamplings. Meanwhile, the QIIME-based literature appears to use
``rarefy'' and ``rarefaction'' interchangeably to mean only a single
subsampling (27, 28). Confusion comes from the WNWN authors' admonition
that ``{[}i{]}n many cases researchers have also failed to repeat the
random subsampling step''. This seems to call on researchers to use
rarefaction rather than the single subsampling step. Subsequent
researchers have continued to conflate the terms when citing WNWN
(\textbf{Supplemental Text}). An exemplar of the confusion is the
creation of a technique that uses ``repeatedly rarefying'' as an
approach distinct from rarefaction when they were in fact re-proposing
traditional rarefaction (29).

@@

The code contained within
\texttt{simulation-cluster-accuracy-server.Rmd} was spilt into
individual R scripts and modified to make the generation of the
simulated communities more modular and scalable. The code for generating
the simulated communities was run using the \texttt{nr-s1} Conda
environment. The ocean and feces parent distributions were generated as
described above. Five variables were altered to simulate each set of
communities. First, eight mixing fractions were used to manipulate the
effect size between the two simulated treatment groups with 40 samples
per group. These were the same effect sizes used in the WNWN study (1.00
{[}i.e., no difference{]}, 1.15, 1.25, 1.50, 1.75, 2.00, 2.50, and
3.50). Second, the number of sequences in each of the 80 samples was
either randomly selected from the 26 library sizes contained within the
GlobalPatterns dataset or from a log distribution of 80 sequencing
depths evenly distributed between the most shallow (58,688 sequences)
and deeply (2,357,181 sequences) sequenced samples in the GlobalPatterns
dataset (Figure S6). Third, the resulting 80 sequencing depths were
scaled so that the median sequencing depth across the 80 samples was
either 1,000, 2,000, 5,000, 10,000, or 50,000 sequences. Fourth, the
sequencing depth of each sample was either randomly assigned to each
treatment group or assigned so that those depths less than or greater
than the median were assigned to separate treatment groups. Finally,
each set of conditions was replicated 100 times. These five parameters
resulted in 1,600 simulated datasets (8 effect sizes x 2 sequencing
depth models x 5 sequencing depths x 2 sequencing depth assignment
models x 100 replicates).

@@

\textbf{Figure 3. When the median sequencing depth was 2,000 sequences
or more, rarefaction of the entire dataset performed better than
removing the smallest 15\% of samples when using K-means clustering.}
This figure is analogous to Figure S4 except that K-means clustering was
used instead of PAM.

\textbf{Figure 4. K-means clustering of distances calculated with
rarefaction were as good or better than any other normalization method.}
This figure is analogous to Figure S3 except that K-means clustering was
used instead of PAM, rarefaction on the full dataset was used instead of
subsampling to the size of the sample at the 15th percentile, and DESeq
Variance Stabilization normalized OTU counts were only used with
Euclidean distances.

\textbf{Figure 5. Clustering accuracies that used rarefaction were as
good or better than the other normalization procedures when a log-scaled
distribution of sequencing depths.} This figure is analogous to Figure 4
except that the sequencing depths for each of the 80 samples in each
simulation were drawn without replacement from a log-scaled distribution
rather than from the GlobalPatterns sequencing depths.

### 2023-11-09 submission/supplemental_text.Rmd
Fix some typos that appeared in proofs

---
bibliography: references.bib
output:
  pdf_document:
    keep_tex: true
csl: asm.csl
geometry: margin=1.0in
header-includes:
 - \usepackage{setspace}
 - \doublespacing
 - \usepackage{helvet} % Helvetica font
 - \renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
 - \usepackage[T1]{fontenc}
 - \usepackage[shortcuts]{extdash}
 - \usepackage[document]{ragged2e}
---

@@

Quotes from selected papers that cited WNWN and conflated "rarefying"/"rarefy" and "rarefaction". Many of these papers also attributed "rarefaction" to WNWN. Rarefying, rarefy, and rarefaction are bolded to make the words easier to identify. All references have been updated to reflect the numbering in the current study. Citations to WNWN are indicated by [@McMurdie2014]. Quotes are ordered chronologically.

@@

>>>>

---
bibliography: references.bib
output: pdf_document
csl: asm.csl
geometry: margin=1.0in
header-includes:
 - \usepackage{setspace}
 - \doublespacing
 - \usepackage{helvet} % Helvetica font
 - \renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
 - \usepackage[T1]{fontenc}
 - \usepackage[shortcuts]{extdash}
 - \usepackage[document]{ragged2e}
---

@@

Quotes from selected papers that cited WNWN and conflated "rarefying"/"rarefy" and "rarefaction". Many of these papers also attributed "rarefaction" to WNWN. Rarefying, rarefy, and rarefaction are bolded to make the words easier to identify. All references have been updated to reflect the numbering in the current study. Citations to WNWN are indicated by [@McMurdie2014] and all numbering is in reference to the list of references in this Supplement. Quotes are ordered chronologically.

@@

\includegraphics{figure_S1.png}

**Figure S1. Re-running the R markdown files provided in Protocol S1 of WNWN qualitatively reproduced Figure 4 from WNWN.**

\includegraphics{figure_S2.png}

**Figure S2. Re-running the R markdown files provided in Protocol S1 of WNWN qualitatively reproduced Figure 5 from WNWN.**

\includegraphics{figure_S3.png}

**Figure S3. Successful reimplementation and expansion of analysis presented in Figure 4 from WNWN in a Snakemake pipeline.** The reimplemented workflow largely borrowed from the original `simulation-cluster-accuracy-server.Rmd` R markdown file provided in WNWN's Protocol S1. The most notable differences include the use of 100 rather than 5 randomizations and the addition of the median sequencing depth (&Ntilde;~L~) of 50,000. The plotting symbols indicate the median of 100 randomizations and the error bars represent the observed 95% confidence interval. Simulations run at the same effect size are dodged to better reveal overlapping data. The sequencing depths were drawn from the GlobalPatterns dataset and sequences were clustered using PAM.

\includegraphics{figure_S4.png}

**Figure S4. Successful reimplementation and expansion of analysis presented in Figure 5 from WNWN in a Snakemake pipeline.** The reimplemented workflow largely borrowed from the original `simulation-cluster-accuracy-server.Rmd` R markdown file provided in WNWN's Protocol S1. The most notable differences include the use of 100 rather than 5 randomizations and the addition of the median sequencing depth (&Ntilde;~L~) of 50,000. The plotting symbols indicate the median of 100 randomizations and the error bars represent the observed 95% confidence interval. Simulations run at the same effect size are dodged to better reveal overlapping data. A light gray line is shown to indicate the best possible accuracy for each library size minimum quantile value. The sequencing depths were drawn from the GlobalPatterns dataset and sequences were clustered using PAM.

\includegraphics{figure_S5.png}

**Figure S5. DESeq Variance Stabilization of OTU counts resulted in negative values that were used to calculate Bray-Curtis and Weighted UniFrac distances.** The median number of negative OTU counts that had a negative OTU count following normalization increased with the effect size and decreased as &Ntilde;~L~ increased (first row). The error bars indicate the observe 95% confidence interval. The fraction simulated datasets that had a negative OTU count following normalization also increased with effect size, but increased as &Ntilde;~L~ increased (second row). For each effect size there were 100 replicate datasets.

\includegraphics{figure_S6.png}

**Figure S6. Comparison of the normally distributed sequencing depths from the GlobalPatterns dataset and a log-scaled distribution of sequencing depths.** The log-scaled distribution was generated so that each sample in a simulation could have a unique number of sequences and to simulate the skew right distribution commonly seen in microbiome studies.

\includegraphics{figure_S7.png}

**Figure S7. Rarefaction with all samples yielded clustering accuracies that were as good or better than removing the smallest 15% of samples across distance calculation methods.** This figure is analogous to Figure 3 except that the sequencing depths for each of the 80 samples in each simulation were drawn without replacement from a log-scaled distribution rather than from the GlobalPatterns sequencing depths.

\includegraphics{figure_S8.png}

**Figure S8. Normalization and distance calculation methods vary in their sensitivity to removal of rare OTUs.** Larger values indicate that the clustering accuracy from filtered datasets were larger than those from non-filtered datasets. The median of 100 randomizations did not meaningfully vary from 0.0, but the observed 95% confidence interval varied considerably. Data are shown for a median sequencing depth (&Ntilde;~L~) of 10,000 sequences when individual sequencing depths were sampled with replacement from the GlobalPatterns dataset or without replacement from the log-scaled distribution.

\includegraphics{figure_S9.png}

**Figure S9. Rarefaction was consistently as good or better than all other normalization methods at controlling for Type I error and maximizing power to detect differences in treatment group using alpha-diversity metrics regardless of whether sequencing depth was confounded by treatment group when using sequencing depths drawn from a lognormal distribution.** Statistical comparisons of OTU richness and Shannon diversity were performed using the non-parametric Wilcoxon two-sampled test. Type I errors were assessed as the fraction of 100 simulations that yielded a significant P-value (i.e., less than or equal to 0.05) at an effect size of 1.00. Power was assessed as the fraction of 100 simulations that yielded a significant P-Value at an effect size of 1.15. Data are shown for when the case when individual sequencing depths were sampled without replacement from the log-scaled distribution.
